---
title: "Graticule Exercise"
author: "GRV"
date: "8/4/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# CODE
  
  
### Load libraries
```{r libraries, include=FALSE, warning = FALSE, message = FALSE}

library(dplyr) # add pipes 
library(MatchIt)
library(readr) # load and read datasets
library(tibble) # format datasets
library(gtsummary)
library(DescTools) # calculate median and median CI

```


## Load the data tables into dataframe(s) or a database

```{r load, include=FALSE, warning = FALSE, message = FALSE}
setwd("/bd-fs-mnt/ishome/home/o993927/test")
pat = read_csv("patient_table.csv")
enc = read_csv("encounter_table.csv")
```


## Create study cohort

```{r create_cohort, include=FALSE, warning = FALSE, message = FALSE}

cohort = enc %>%
  arrange(patientid, encounterdate) %>%
  group_by(patientid) %>%
  filter(encounterdate > as.Date("2022-07-01") & encounterdate < as.Date("2022-12-31")) %>%
  left_join(pat, by = "patientid") %>%
  mutate(age = as.numeric(difftime(encounterdate,birthdate, units = "weeks"))/52.25)

```  


## Create exposure cohort from July 1st though Dec 31st 2022 for both cohort selection and outcome follow-up
### (Exposure enters a subject into the exposed cohort and serves as the index date)    
#### Keep the first exposure if multiple exposures in same individual 

```{r expose_coh, include=FALSE, warning = FALSE, message = FALSE}

expo_coh = cohort %>% 
  group_by(patientid) %>%
  filter(exposure == TRUE) %>%
  distinct(patientid, .keep_all= TRUE) %>%
  add_column(exp_cohort_indic = 1) 

print(paste0("Exposure cohort length:", " ",dim(expo_coh)[1]))

```

## Explore mortality on merged cohort
```{r check_mortality, include=FALSE, warning = FALSE, message = FALSE}

which(expo_coh$outcome == TRUE & !is.na(expo_coh$deceaseddate))
expo_coh[dput(which(expo_coh$outcome == TRUE & !is.na(expo_coh$deceaseddate))),] # NONE
expo_coh[dput(which(expo_coh$exposure == TRUE & !is.na(expo_coh$deceaseddate))),] # NO NEED TO REMOVE
expo_coh[dput(which(!is.na(expo_coh$deceaseddate))),] # NO NEED TO REMOVE

```


## Compute median age at index date for the exposed cohort (including CI at 95% confidence level)

```{r median_age, include=FALSE, warning = FALSE, message = FALSE}

MedianCI(expo_coh$age, conf.level=0.95)

```

## 3.	Compute aggregate time at risk for the exposed cohort, in person-days  
  
```{r person_days, include=FALSE, warning = FALSE, message = FALSE}


person_day_expo = cohort %>%
  filter(any(exposure == TRUE)) %>%
  mutate(outcm = ifelse(outcome == FALSE, 0, 1)) %>%
 filter(cummax(outcm) <1) %>%
  mutate(duration = encounterdate - first(encounterdate)) %>%
  slice(n()) %>%
  ungroup() %>%
  summarise(person_days = sum(duration))

person_day_expo
   
```


## Write pseudo-code to select a 1:1 unexposed cohort based on the method outlined above
### •	Exposed subjects are exact matched with unexposed subjects based on age, gender, and a healthcare encounter within the same week  
### •	Matching is performed without replacement  
### •	Exposed subjects may serve as unexposed matches during any week before the week of their exposure date  

```{r match_unexposed_pseudocode, include=FALSE, warning = FALSE, message = FALSE}

# prepare exposure cohort for merge with full cohort to get exposure indicators
LOAD data set "expo_coh"
  KEEP variables patientid & exp_cohort_indic
  SAVE as modified variables as "exposed"
END
  

# load and prepare cohort
LOAD data set "cohort" 
  CREATE new column "enc_year_week" 
  MERGE data set "exposde" by patientid
    SAVE new data as "cohort_expose_flag"
END
  

# create categories for matching
WITH data set "cohort_expose_flag"
  FILTER IN exposed
  ORDER BY age, gender & enc_year_week
  SELECT distinct age, gender & enc_year_week
      SUBSET independent categories (factorials)
        SELECT age, gender, enc_year_week & subset 
          SAVE data as "match_factors"
END 

# create exposed data with matching categories
WITH "cohort_expose_flag"
  FILTER IN exp_cohort_indic flag 
    MERGE to "match_factors" by patient_id, age, gender & enc_year_week
      SAVE "expo_factors"
END

# create unexposed data with matching categories
WITH "cohort_expose_flag"
  KEEP all encounters up to exposure week (exlcuding)
    MERGE to "match_factors" by patient_id, age, gender & enc_year_week
      SAVE "unexpo_factors"
END  

WITH "unexpo_factors" and "expo_factors"
  TEST balance of variables
END


# exposed and unexposed
BIND "expo_factors" with "unexpo_factors"
  SELECT one unexposed match for every exposed 
    SAVE "match"
END

WITH "match"
  TEST balance of matched variables
END
  

```



